---
title: "Visualize Filtering"
author: ""
date: "`r Sys.Date()`"
output: 
  html_document:
    self_contained: TRUE
    code_folding: "hide"
    toc: true
    toc_float: true
    toc_collapsed: true
toc_depth: 3
number_sections: true
theme: lumen
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 72, dev = "jpeg")
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      all_times[[options$label]] <<- round(lubridate::as.duration(res))
    }
  }
}))
knitr::opts_chunk$set(time_it = TRUE)
```
# Setup

load packages

```{r setup_env}
suppressPackageStartupMessages({
  library(magrittr)
  library(scater)
  library(ggplot2)
  library(SpatialExperiment)
})
```

```{r}
sce <- readRDS(here::here("results/spe/raw/SPE_combined_markedfilt.rds"))
sce <- readRDS(here::here(snakemake@input[["spe"]]))
```

# Quality control

## Flag border cells

```{r}
table(sce[["is_border_cell"]], useNA="ifany")
```

## image size

```{r}
table(sce[["is_small_image"]], useNA="ifany")
```

## cell size

flag cells with NA area

```{r}
table(sce[["is_na_area"]], useNA="ifany")
```

flag cells with area smaller than `r snakemake@params[["min_cell_area"]]` pixels
```{r}
table(sce[["is_small_cell"]], useNA="ifany")
```


## celltype markers

flag cells with mean cell type marker expression == 0 (all are equal to zero)

```{r}
table(sce[["is_zero_celltype_expression"]], useNA="ifany")
```




## number of cells per image

```{r, fig.height=10,fig.width=10}
ncells <- colData(sce[,sce$to_keep]) %>%
    as.data.frame() %>% 
    dplyr::group_by(sample_id) %>% 
  dplyr::summarise(cell_area = sum(area),
                   n=dplyr::n(),
                   no_pixels=mean(no_pixels),
                   covered_area = cell_area / no_pixels)

table(ncells$n<100,useNA="ifany")

fit <- lm(sqrt(n)~no_pixels,ncells)
# plot(fit)
sfit <- summary(fit)
# which.max(abs(residuals(sfit)))


pltlabels <- ncells[order(abs(residuals(sfit)),decreasing = "TRUE")[seq_len(min(length(unique(sce$sample_id)),5))],] 

ggplot(ncells,aes(no_pixels,sqrt(n),color=covered_area)) +
  geom_point() +
  # geom_smooth(method = "lm",formula = y~x) +
  geom_abline(intercept = coef(sfit)[1,1], slope=coef(sfit)[2,1]) +
  ggrepel::geom_label_repel(aes(no_pixels,sqrt(n),label=sample_id),data=pltlabels)
```




## area coverage

```{r}
img_to_cytoimagelist <- function(img, mask, sample_name){
  stopifnot(is(img, "Image"))
  stopifnot(is(mask, "Image"))
  stopifnot(is(sample_name, "character"))
  
  # create CytoImageList object
  cytoimg <- cytomapper::CytoImageList(img)
  # add metadata
  cytomapper::channelNames(cytoimg) <- dimnames(img)[[3]]
  names(cytoimg) <- as.character(sample_name)
  mcols(cytoimg) <- DataFrame(sample = sample_name ,
                                   sample_id = sample_name)
  
  # Create cellmask
  tmpmask <- as.matrix(EBImage::imageData(mask))
  # convert to integers
  tmpmask <- matrix(as.integer(tmpmask), nrow = dim(tmpmask)[1], ncol = dim(tmpmask)[2])
  all(is.integer(tmpmask))
  all(is.integer(EBImage::imageData(mask)))
  EBImage::imageData(mask) <- tmpmask
  # create CytoImageList object for cell mask
  cytomask <- cytomapper::CytoImageList(mask)
  # add metadata
  names(cytomask) <- as.character(sample_name)
  mcols(cytomask) <- DataFrame(sample = sample_name ,
                               sample_id = sample_name)
  return(list(cytoimg=cytoimg, cytomask=cytomask))
}
```

```{r, fig.width=20}
area_coverage <- colData(sce[,sce$to_keep]) %>%
    as.data.frame() %>%
    dplyr::group_by(sample_id) %>%
    dplyr::summarize(cell_area = sum(area),
           no_pixels = mean(width_px) * mean(height_px)) %>%
    dplyr::mutate(covered_area = cell_area / no_pixels) 


# check for outlier in area coverage
immad <- mad(na.omit(area_coverage$covered_area))
immed <- median(na.omit(area_coverage$covered_area))

area_coverage %>%
    ggplot() +
        geom_point(aes(sample_id, covered_area)) + 
        theme_minimal(base_size = 15) +
        ylim(c(0, 1)) + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8)) +
        ylab("% covered area") + xlab("") +
  geom_abline(slope=0,intercept=immed)+
  geom_abline(slope=0,intercept=immed+2*immad)+
  geom_abline(slope=0,intercept=immed-2*immad)

is_outlier <- (area_coverage$covered_area < (immed-2*immad)) | (area_coverage$covered_area > (immed+2*immad))
area_coverage[is_outlier,c("sample_id","cell_area","no_pixels","covered_area")]
```


```{r cytomapper_plotPixels, fig.width=10, fig.height=10, results='asis'}
cat("### Outliers {.tabset}\n\n")
for(i in which(is_outlier)){
  cat(paste0("#### ",area_coverage[i,][["sample_id"]],"\n\n"))
  samind <- area_coverage$sample_id[[i]]
  img_file_name <- snakemake@input[["images"]][stringr::str_detect(snakemake@input[["images"]],samind)]
  img <- EBImage::readImage(here::here(img_file_name))
  mask_file_name <- snakemake@input[["masks"]][stringr::str_detect(snakemake@input[["masks"]],samind)]
  mask <- EBImage::readImage(here::here(mask_file_name))*(2^16-1)
  
  cytols <- img_to_cytoimagelist(img,mask, area_coverage$sample_id[i])
  
  tmpsce <- sce[,sce$sample_id == area_coverage$sample_id[i]]
  hist(colMeans(assay(tmpsce,"exprs")[c("DNA1","DNA2"),]),breaks=100,
       main="Histogram of DNA intensity", xlab="Mean DNA intensitiy")
  cat("\n\n")
  cytomapper::plotPixels(cytols$cytoimg, tmpsce,mask = cytols$cytomask, cell_id = "ObjectNumber", img_id = "sample_id", colour_by = as.character(which(rownames(tmpsce)=="DNA1")),legend=NULL, colour = list(`36`=c("black",viridis::viridis(2)[2])))
  cat("\n\n")
}
```


# Flag outlier cells (marker intensity)

```{r}
table(sce[["is_outlier_expression"]])
```

# Summary

Cells to remove:

```{r}
table(!sce[["to_keep"]])
```


